Machine Learning for Predictive Auto-Tuning with Boosted Regression Trees

### 2011-12-14 paper/draft1.tex

\begin{figure*}[!t]
\centering
\includegraphics[width=.9\linewidth]{illustrator/fig_main_R1.pdf}
\caption{
Computation speed for novel problem configurations using predictive
vs. empirical auto-tuning.
(a) As more problem configurations are used to train the regression model,
the performance of predictive auto-tuning approaches that of empirical auto-tuning
despite taking only 0.1 seconds instead of minutes.
(b) and (c) For a wide range of devices, implementation speeds found by predictive auto-tuning correlate tightly with
implementation speeds found by empirical auto-tuning.
%The boosted regression trees used for predictive auto-tuning generalize correctly across the implementation space.
}
\label{fig:fig_gflop_scatter}
\end{figure*}

Figure~\ref{fig:allstars} shows the effectiveness of empirical auto-tuning is
in this setting.  Taking the GTX 580 as an example, and averaging across the
range of problem configurations in our study, we find that empirically
auto-tuned implementations are on average about 50\% faster than the reference
implementation.  The reference in turn, is about 50\% faster than
implementations that were empirically auto-tuned for a \emph{randomly chosen}
different argument configuration.  This shows that it is genereally not enough
to auto-tune for particular argument configurations; instead it is important to
choose the right kernel for the job for each unique argument configuration
(input-dependent auto-tuning).

@@

Figure~\ref{fig:fig_gflop_scatter} (b) and (c) show that the implementation
speeds found by predictive auto-tuning correlates very well with the speed of
empirical auto-tuning across all of the devices we tested: GeForce GTX 295,
Tesla C1060, GeForce GTX 580, GeForce GTX 480 and Tesla C2070 which
span two generations of NVIDIA CUDA platforms.

% backtracking
In some cases, predictive auto-tuning yields an invalid implementation.  We
dealt with this scenario by backtracking through the various best-estimates
found during the hill-climbing search.  Some invalid kernels can only be
discovered after compiling code and attempting to run the compiled code, so in
these cases predictive auto-tuning took up to 3 seconds.  Even in these cases
the predictive model is much faster than empirical auto-tuning, because the
first kernel that actually runs is a fast one.

Figure~\ref{fig:fig_ntrain} shows how the training set size
and the value of $\zeta$ affect the accuracy of predictive auto-tuning.
All candidates timed during the grid and
hill-climbing search procedures were used as training examples, so the training
set sizes ranged from an average of 1,480 (10 problem configurations) to 29,600
(200 problem configurations).
Training from the largest training sets took approximately 30 seconds.
Training 50 or more problem configurations yielded quite accurate predictions
with $\zeta=\log(0.5)$, which was the best value for $\zeta$ across the range of
devices in our study.

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{illustrator/fig_allstars_mixup_580_R1.pdf}
\caption{Different arguments call for different kernels:
{\em left} is the average argument-specific empirical
auto-tuned performance across 100 random argument configurations,
{\em middle} is the average speed of our reference implementation,
{\em right} is the average speed of kernels auto-tuned for different argument
configurations than the one being tested.  For 38 out of 100 random configurations, the
kernel auto-tuned for another problem could not even run on the GTX 295
hardware.  These points contributed a speed of 0, bringing down the average
much lower than the reference.  Good performance across a variety of inputs
requires {\em input-dependent auto-tuning}.
}
\label{fig:allstars}
\end{figure}

>>>>

\begin{figure*}[!t]
\centering
\includegraphics[width=.9\linewidth]{illustrator/fig_main_R1.pdf}
\caption{
Computation speed for novel problem configurations using predictive
vs. empirical auto-tuning.
(a) As more problem configurations are used to train the regression model,
the performance of predictive auto-tuning approaches that of empirical auto-tuning
despite taking only 0.1 seconds instead of 2 minutes.
(b) and (c) For a wide range of devices, implementation speeds found by predictive auto-tuning correlate tightly with
implementation speeds found by empirical auto-tuning.
%The boosted regression trees used for predictive auto-tuning generalize correctly across the implementation space.
}
\label{fig:fig_gflop_scatter}
\end{figure*}

Figure~\ref{fig:allstars} shows the effectiveness of empirical auto-tuning is
in this setting.  Taking the GTX 295 as an example, and averaging across the
range of problem configurations in our study, we find that empirically
auto-tuned implementations are on average about 50\% faster than the reference
implementation.  The reference in turn, is about 50\% faster than
implementations that were empirically auto-tuned for a \emph{randomly chosen}
different argument configuration.  This shows that it is genereally not enough
to auto-tune for particular argument configurations; instead it is important to
choose the right kernel for the job for each unique argument configuration
(input-dependent auto-tuning).

@@

Figure~\ref{fig:fig_gflop_scatter} (b) and (c) show that the implementation
speeds found by predictive auto-tuning correlates very well with the speed of
empirical auto-tuning across all of the devices we tested: GTX 580, GTX 480,
GTX 295, Tesla C2070, and Tesla C1060 which span three generations of
CUDA platforms.

% backtracking
In some cases, predictive auto-tuning yields an invalid implementation.  We
dealt with this scenario by backtracking through the various best-estimates
discovered during the hill-climbing search.  Some invalid kernels can only be
discovered after compiling code and attempting to run the compiled code, so in
these cases predictive auto-tuning took up to 3 seconds.  Even in these cases
the predictive model is much faster than empirical auto-tuning, because the
first kernel that actually runs is a fast one.

Figure~\ref{fig:fig_ntrain} shows how the training set size
and the value of $\zeta$ affect the accuracy of predictive auto-tuning.
All candidates timed during the grid and
hill-climbing search procedures were used as training examples, so the training
set sizes ranged from an average of 1480 (10 problem configurations) to 29 600
(200 problem configurations).
Training from the largest training sets took approximately 30 seconds.
Training 50 or more problem configurations yielded quite accurate predictions
with $\zeta=\log(0.5)$, which was the best value for $\zeta$ across the range of
devices in our study.

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{illustrator/fig_allstars_mixup_580_R1.pdf}
\caption{Different arguments call for different kernels:
{\em left} is the average argument-specific empirical
auto-tuned performance across 100 random argument configurations,
{\em middle} is the average speed of our reference implementation,
{\em right} is the average speed of kernels auto-tuned for different argument
configurations than the one being tested.  For 38/100 random configurations, the
kernel auto-tuned for another problem could not even run on the GTX 295
hardware.  These points contributed a speed of 0, bringing down the average
much lower than the reference.  Good performance across a variety of inputs
requires {\em input-dependent auto-tuning}.
}
\label{fig:allstars}
\end{figure}
